Sketching for Density Estimation

### 2018-11-12 density.tex
push mid flight

\section{General framework for sketching}
In many sketching problems our goal is to approximate a function of the of the form $F(q) = \sum_{i=1}^{n} f(x_i, q)$ where $x_i$ are all the stream items and and $q$ is some query point. 
The goal is to produce a sketch of $\ell$ points $z$ and weights $w$ such that $\tilde F(q) = \sum_{i=1}^{\ell}w_i f(z_i,q) \approx \sum_{i=1}^{n} f(x_i, q)$. 
If $z \subset x$ this is a subset selection problem. 
For bounded functions $f$, uniform sampling combined with a union bound over the space of possbile values for $q$ always provides a valid solution.
However, it is often far from being the best possible in terms of its space-accuracy tradeoff.

Here are some examples:
\begin{itemize}
\item in approximate counting we have $f(x, q) = 1$ if $x=q$ and zero else.
\item in quantile approximation we have $f(x, q) = 1$ if $x<q$ and zero else.
\item in matrix row subset selection we have $f(x, q) = \langle x,q \rangle ^2$.
\item in density estimation $f(x, q) = \exp(- \|x-y\|_2^2/\sigma^2)$ or any other kernel (see above)
\item in machine learning $f(x, q) = \operatorname{loss}(x,q)$ where $q$ is the parameters of the model and $x$ is a training example and its label.
\end{itemize}
For all of the above, we can find flat error functions $E(q) = \sum_{i=1}^{k} s_i f(x_i,q)$ such that $\sup_q E(q)$ is small. 
We argue that this gives a general mechanism for sketching in a diverse set of scenarios.

>>>>

\section{General framework for sketching}
In many sketching problems our goal is to approximate a function of the of the form $F(q) = \sum_{i=1}^{n} f(x_i, q)$ where $x_i$ are all the stream items and and $q$ is some query point. 
The goal is to produce a sketch of $\ell$ points $z$ and weights $w$ such that $\tilde F(q) = \sum_{i=1}^{\ell}w_i f(z_i,q) \approx \sum_{i=1}^{n} f(x_i, q)$. 
If $z \subset x$ this is a subset selection problem. 
For bounded functions $f$, uniform sampling combined with a union bound over the space of possible values for $q$ always provides a valid solution. 
While the number of possible queries $q$ is often infinite it often reduces to a finite set through a net argument. 
This is common practice in machine learning, specifically, PAC learning.

However, it is often far from being the best possible in terms of its space-accuracy tradeoff.

Here are some examples:
\begin{itemize}
\item in approximate counting we have $f(x, q) = 1$ if $x=q$ and zero else. Here both $x$ and $q$ belong to some finite domain.
\item in quantile approximation we have $f(x, q) = 1$ if $x<q$ and zero else. Here, $x$ and $q$ belong to a set which exhibits strong ordering.
\item in matrix row subset selection we have $f(x, q) = \langle x,q \rangle ^2$ where $q$ is any unit vector in $\R^d$ and $x$ are matrix rows. For simplicity, we will assume in this manuscript that $x$ is also unit norm.
\item in density estimation $f(x, q) = \exp(- \|x-y\|_2^2/\sigma^2)$ or any other kernel (see \cite{})
\item in logistic regression $f(x, q) = \operatorname{loss}(x,q) = 1/(1 + \exp(-q^T x))$. 
Here $q$ is the parameters of the linear model and $x$ is a training example vector multiplied by its label ($1$ or $-1$).
\item Quadratic forms of the graph Laplacian are $f(x, q) = q^T L_x q$ where $x$ is an edge and $L_x$ is the graph laplacian and $q$ is a test vector.
\item Probably also good for matrix approx when we get entry updates. 
\end{itemize}

We argue that we can offer a unified solution for producing coresets for all of the above problems in a unified manner. 
Moreover, our solution created streaming algorithms with fully mergeable sketches.

We begin with describing the basic operation of compaction and the notion of flat error functions.

Let us consider a signed sum of error function $E(q) = \sum_{i=1}^{n} s_i f(x_i,q)$.
Now consider $\tilde F(q) = F(q) + E(q) = \sum_{i=1}^{n} f(x_i,q)  + \sum_{i=1}^{n} s_i f(x_i,q)  = 2 \sum_{i | s_i=1} f(x_i,q)$.
Therefore $\tilde F(q)$ is a coreset for $F(q)$ with item weights of $2$ and error at most $|\tilde F(q) - F(q)| \le \max_q | E(q)|$.
Note that the above is true for any choice of signs $s_i$, specifically, for those minimizing $\max_q | E(q)|$.
We call $E$ flat if $\min_s \max_q |E(q)| = o(n)$.

\begin{definition}
We define $E_n$ with respect to a function $f(x,q)$ to be  $\min_s \max_q \max_x \sum_{i=1}^{n} s_i f(x_i,q)$. 
\end{definition}

As a warmup exercise, consider the case where $E_n = 1$.
This is the case for quantiles and for counting. 
At step one, we will create a coreset of weight $2$ and size at most $n/2$ (by either adding or subtracting $E$). Here we incur an error of at most $1$.
Then, we repeat the process and incur an error of at most $2$ resulting in a stream of length at most $n/4$. 
We can continue the above process until we are left with $O(1/\eps)$ points in the coreset. At this point, the error adds up to at most $\eps n$.
Similarly, when $E_n = O(\sqrt{n})$ we get a coreset of size $O(1/\eps^2)$ whose error is at most $\eps n$.
This gives an offline algorithm.

\zk{We can describe the general streaming algorithm here. I think $E_n$ is a sufficient characterization of this.}

For all of the above, we argue that we can find flat signed error functions such that $E_n = O(1)$ or $E_n = O(\sqrt{n})$.
